---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qgisprocess

<!-- badges: start -->
[![R build status](https://github.com/paleolimbot/qgisprocess/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/qgisprocess/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of `qgisprocess` is to provide an R interface to the popular and open source desktop geographic information system (GIS) program [QGIS](https://qgis.org/en/site/).
The package is a re-implementation of functionality provided by the archived [RQGIS](https://cran.r-project.org/package=RQGIS) package, which was partially revived in the [RQGIS3](https://github.com/r-spatial/RQGIS3) package.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/qgisprocess")
```

The qgisprocess package wraps the `qgis_process` command-line utility, which is available in QGIS >= [3.14.16](https://github.com/qgis/QGIS/releases/tag/final-3_14_16), [released](https://qgis.org/en/site/getinvolved/development/roadmap.html) in September 2020. MacOS users will have to install the a recent nightly build until QGIS 3.16 has been released. Download instructions for all platforms are available at <https://download.qgis.org/>.

## Example

The following example demonstrates the [buffer](https://docs.qgis.org/testing/en/docs/user_manual/processing_algs/qgis/vectorgeometry.html#buffer) algorithm in action. The passing of [sf](https://r-spatial.github.io/sf) and [raster](https://cran.r-project.org/package=raster) objects is experimentally supported (and will be well-supported in the future!).

```{r buffer}
library(qgisprocess)
input <- sf::read_sf(system.file("shape/nc.shp", package = "sf"))
output_file <- file.path(tempdir(), "nc_buffered.gpkg")

qgis_run_algorithm(
  "native:buffer",
  INPUT = input,
  DISTANCE = 1,
  SEGMENTS = 10,
  DISSOLVE = 'True',
  END_CAP_STYLE = 0,
  JOIN_STYLE = 0,
  MITER_LIMIT = 10,
  OUTPUT = output_file
)
output_sf <- sf::read_sf(output_file)
unlink(output_file)

plot(sf::st_geometry(output_sf))
```

You can read the help associated with an algorithm using `qgis_show_help()`. It may also be useful to run an algorithm in the QGIS GUI and examine the console output to determine how the various input values are translated to processing arguments.

```{r}
qgis_show_help("native:buffer")
```

## Further reading

- A [paper](https://journal.r-project.org/archive/2017/RJ-2017-067/index.html) on the original RQGIS package published in the R Journal
- A [discussion](https://github.com/r-spatial/discuss/issues/41) options for running QGIS from R that led to this package
- The [pull request](https://github.com/qgis/QGIS/pull/34617) in the QGIS repo that led to the development of the `qgis_process` command-line utility 
