---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qgisprocess

<!-- badges: start -->
[![R build status](https://github.com/paleolimbot/qgisprocess/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/qgisprocess/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of `qgisprocess` is to provide an R interface to the popular and open source desktop geographic information system (GIS) program [QGIS](https://qgis.org/en/site/).
The package is a re-implementation of functionality provided by the archived [`RQGIS`](https://cran.r-project.org/package=RQGIS) package, which was partially revived in the [`RQGIS3`](https://github.com/r-spatial/RQGIS3) package.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/qgisprocess")
```

## Installing QGIS on your computer

The `qgis_process` command-line utility is available in QGIS >= [3.14.16](https://github.com/qgis/QGIS/releases/tag/final-3_14_16), [released](https://qgis.org/en/site/getinvolved/development/roadmap.html) in September 2020.
You can install this version of QGIS on Linux, Mac and (via the OSGeo4W package) Windows, as described at [qgis.org](https://qgis.org/en/site/forusers/download.html).

## qgisprocessing in Docker

You can also use Docker images containing the necessary dependencies to run this package in a container.
The [`geocompr/geocompr:qgis`](https://github.com/geocompr/docker) image, for example, has QGIS and RStudio preinstalled for interactive use and can be downloaded from Dockerhub and run as follows:

```
docker run -d -p 8788:8787 -e USERID=$UID -e PASSWORD=ps -v ${pwd}:/home/rstudio/ geocompr/geocompr:qgis
```

After running this command (having set a more secure password as appropriate) you should be able to access RStudio Server from http://localhost:8788/.
Once inside you can install `qgisprocess` as documented here.

You can run also `qgisprocessing` in a Dockerfile provided in this repo as follows:

```
# do this once:
# docker build . --tag qgisprocess-devel

# do this to run an interactive R session in the docker image
docker run --rm -it -v $(pwd):/qgisprocess -w /qgisprocess qgisprocess-devel R
```

Once in the container, you can use `devtools::load_all()`, `devtools::test()`, and `devtools::check()` to develop the package.


## Examples

This is a basic example which tests that the package can detect a working version of `qgis_process`:

```{r example}
library(qgisprocess)

# basic QGIS info
qgis_path()
qgis_version()
qgis_algorithms()

# get help
qgis_show_help("native:filedownloader")

# run the algorithm!
qgis_run_algorithm(
  "native:filedownloader", 
  URL = "https://httpbin.org/get",
  OUTPUT = "test-file.json"
)
```

```{r, echo=FALSE}
unlink("test-file.json")
```

The following examples demonstrates the [buffer](https://docs.qgis.org/testing/en/docs/user_manual/processing_algs/qgis/vectorgeometry.html#buffer) algorithm in action.

```{r buffer}
input_file <- sf::read_sf(system.file("shape/nc.shp", package = "sf"))
output_file <- file.path(tempdir(), "nc_buffered.gpkg")
qgis_run_algorithm(
  "native:buffer",
  INPUT = input_file,
  DISTANCE = 1,
  SEGMENTS = 10,
  DISSOLVE = 'True',
  END_CAP_STYLE = 0,
  JOIN_STYLE = 0,
  MITER_LIMIT = 10,
  OUTPUT = output_file
)
output_sf <- sf::read_sf(output_file)
sf::st_crs(output_sf)
plot(sf::st_geometry(output_sf))
```


## Further reading

- A [paper](https://journal.r-project.org/archive/2017/RJ-2017-067/index.html) on the original RQGIS package published in the R Journal
- A [discussion](https://github.com/r-spatial/discuss/issues/41) options for running QGIS from R that led to this package
- The [pull request](https://github.com/qgis/QGIS/pull/34617) in the QGIS repo that led to the development of the `qgis_process` command-line utility 

