---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qgisprocess

<!-- badges: start -->
[![R build status](https://github.com/paleolimbot/qgisprocess/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/qgisprocess/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of qgisprocess is to ...

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/qgisprocess")
```

The `qgis_process` command-line utility will be available in the next release of QGIS; for now you will need a development version of QGIS installed. This is not trivial, but a Dockerfile is provided as a developmet environment.

```
# do this once:
# docker build . --tag qgisprocess-devel

# do this to run an interactive R session in the docker image
docker run --rm -it -v $(pwd):/qgisprocess -w /qgisprocess qgisprocess-devel R
```

Once in the container, you can use `devtools::load_all()`, `devtools::test()`, and `devtools::check()` to develop the package.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(qgisprocess)

# basic QGIS info
qgis_path()
qgis_version()
qgis_algorithms()
```
