---
title: "Intro: Basic usage of qgisprocess"
author: "Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro: Basic usage of qgisprocess}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**qgisprocess** is basically a wrapper around the standalone command line tool [`qgis_process`](https://github.com/qgis/QGIS-Documentation/issues/5803).
Therefore, it already tries to detect where `qgis_process` is installed on your system when it is being attached.

```{r}
library("qgisprocess")
```

If you want to find out, how **qgisprocess**  detected `qgis_process`, run `qgis_configure()`.

```{r config}
qgis_configure()
```

Here, `qgis_process` was detected in the PATH variable.
However, `qgis_configure()` first tries if the R option `qgisprocess.path` or the the global environment variable `R_QGISPROCESS_PATH` has been set.
This already indicates that you can specify one of these variables in case `qgis_process` has not been installed in one of the most common locations which might sometimes happen under Windows.
If `qgis_process` is installed, you can set `options(qgisprocess.path = '/path/to/qgis_process')` and re-run `qgis_configure()`.
In the case of Windows, make sure to indicate the path to the `qgis_process-qgis.bat` file, e.g.,

```{r win-config, eval=FALSE}
# specify path to QGIS installation on Windows 
options("qgisprocess.path" = "C:/Program Files/QGIS 3.14/bin/qgis_process-qgis.bat")
qgis_configure()
```


First, let's check which QGIS version our system is running:

```{r vers}
qgis_version()
```

Next, let's list all available providers including available 3rd party applications: 

```{r providers}
qgis_providers()
```

List all geoalgorithms at your disposal:

```{r algs}
qgis_algorithms()
```

Find out what a specific geoalgorithm is doing, e.g., `native:buffer`.

```{r desc}
qgis_show_help("native:buffer")
# cat(qgis_description("native:buffer"))
# cat(qgis_help_text("native:buffer"))
```

Find out the arguments of a specific geoalgorithm:

```{r args}
qgis_arguments("native:buffer")
```


Vector example

If you would like to have an R-like experience running QGIS, create R functions first.

```{r function-creation}
qgis_buffer = qgis_function("native:buffer")
```


Raster example (GRASS7)

```{r}
qgis_arguments("grass7:r.slope.aspect")
```

```{r slopeaspect}
library("raster")
data("dem", package = "spDataLarge")
projection(dem) = sf::st_crs(32717)$proj4string

info = qgis_run_algorithm(alg = "grass7:r.slope.aspect",
                elevation = dem, 
                #pcurvature = file.path(tempdir(), "pcurv.tif"),
                #tcurvature = file.path(tempdir(), "tcurv.tif"),
                .quiet = TRUE)
raster(qgis_output(info, "pcurvature")[1])
```

<!--
# raster example
## ----slope_aspect, eval = TRUE, messages = FALSE, cache = FALSE----------
params$elevation <- dem
params$pcurvature <- file.path(tempdir(), "pcurv.tif")
params$tcurvature <- file.path(tempdir(), "tcurv.tif")
out <- run_qgis(alg = "grass7:r.slope.aspect",
                params = params,
                load_output = TRUE,
                show_output_paths = FALSE)
class(out)
names(out)

## ----eval = FALSE--------------------------------------------------------
library("raster")
plot(stack(out))

-->
