---
title: "Basic usage of qgisprocess"
author: "Jannes Muenchow & Floris Vanderhaeghe"
date: "Last updated: yyyy-mm-dd"
output:
  html_document: 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Basic usage of qgisprocess}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setting up the system

**qgisprocess** is basically a wrapper around the standalone command line tool [`qgis_process`](https://github.com/qgis/QGIS-Documentation/issues/5803).
Therefore, you need to have installed QGIS on your system as well as third-party providers such as GRASS and SAGA to access and run all geoalgorithms provided through `qgis_process` from within R.

To facilitate using **qgisprocess**, we have created a docker image that already comes with the needed software packages.
You can pull it from Github's container registry by running:

```{sh, eval=FALSE}
docker pull ghcr.io/geocompx/docker:qgis
```

For a more detailed introduction on how to get started with docker, please refer to https://github.com/geocompx/docker.

# Package configuration

Since **qgisprocess** depends on the command line tool `qgis_process`, it already tries to detect `qgis_process` on your system when it is being loaded, and complains if it cannot find it.

```{r}
library("qgisprocess")
```

In fact, once **qgisprocess** has detected `qgis_process`, it will cache among others the path to the executable, the QGIS version and the list of known algorithms.
Therefore, usually you don't have to do any configuration yourself, unless there's a message telling you to do so.

If you are interested in the details about this process, e.g. how **qgisprocess**  detected `qgis_process`, run `qgis_configure(use_cached_data = TRUE)`.

```{r config}
qgis_configure(use_cached_data = TRUE)
```

If needed the cache will be rebuilt automatically upon loading the package.
This is the case when the QGIS version or the location of the `qgis_process` command-line utility has changed, user-settings (e.g. the option `qgisprocess.path`) have been altered or a changed state of the processing provider plugins (enabled vs. disabled) has been detected.

Rebuilding the cache can be triggered manually by running `qgis_configure()` (its default is `use_cached_data = FALSE`).

To determine the location of `qgis_process`, `qgis_configure()` first checks if the R option `qgisprocess.path` or the global environment variable `R_QGISPROCESS_PATH` has been set.
This already indicates that you can specify one of these settings in case `qgis_process` has not been installed in one of the most common locations or if there are multiple QGIS versions available.
If this is the case, set `options(qgisprocess.path = '/path/to/qgis_process')` or set the environment variable (e.g. in `.Renviron`) and run `qgis_configure()`.
Under Windows make sure to indicate the path to the `qgis_process-qgis.bat` file, e.g.,

```{r win-config, eval=FALSE}
# specify path to QGIS installation on Windows 
options(qgisprocess.path = "C:/Program Files/QGIS 3.28/bin/qgis_process-qgis.bat")
# or use the QGIS nightly version (if installed via OSGeo4W)
# options(qgisprocess.path = "C:/OSGeo4W64/bin/qgis_process-qgis-dev.bat")
qgis_configure() # or use library(qgisprocess) if package was not loaded yet
```

Assuming that package loading or `qgis_configure()` ran successfully, we can check which QGIS version our system is running (it takes this from the cache):

```{r vers}
qgis_version()
```

Next, let's list all available providers including available third-party applications: 

```{r providers}
qgis_providers()
```

This tells us that we can also use the third-party providers GDAL, SAGA and GRASS through the QGIS interface.

# Basic usage

## First example

To find out about the available (cached) geoalgorithms, run:

```{r algs}
algs <- qgis_algorithms()
algs
```

Since we have also installed GRASS and SAGA, nearly 1000 geoalgorithms are at our disposal.
To find out about a specific geoalgorithm and a description of its arguments, use `qgis_show_help()`, e.g.:

```{r help, eval=FALSE}
qgis_show_help("native:buffer")
## Buffer (native:buffer)
## 
## ----------------
## Description
## ----------------
## This algorithm computes a buffer area for all the features in an input layer, using a fixed or dynamic distance.
## 
## The segments parameter controls the number of line segments to use to approximate a quarter circle when creating rounded offsets.
## 
##...
```

To find out the arguments of a specific geoalgorithm, run:

```{r args-buffer}
qgis_arguments("native:buffer")
```

And finally run it with `qgis_run_algorithm()`:

```{r buffer}
# if needed, first install spDataLarge:
# remotes::install_github("Nowosad/spDataLarge")
data("random_points", package = "spDataLarge")
out <- qgis_run_algorithm("native:buffer", INPUT = random_points, DISTANCE = 50)
```

As a convenience to the user, `qgis_run_algorithm()` reports all unspecified and automatically chosen arguments.
If you want to have even more information on what is going on in the background, set `.quiet` to `FALSE`.
The `out` object is of class `qgis_result` and contains the path to the output object created by `qgis_process` (when not explicitly setting an output object, **qgisprocess** creates it automatically for you), the used algorithm and input arguments as well as an object `.processx_result` which is especially useful if somethings went wrong.

For QGIS 3.24 and later, `qgis_run_algorithm()` passes the input arguments to QGIS as a JSON string.
The JSON input string is also included in the `out` object.
Moreover, the user can specify input arguments directly as JSON in `qgis_run_algorithm()`.
That is useful since input parameters can be copied from the QGIS GUI as JSON.
This will be demonstrated in a separate tutorial.

```{r inspect}
# inspect the out object
class(out)
str(out)
```

To read in the QGIS output and visualize it, we can run:

```{r plot-buffer, message=FALSE}
library("sf")
library("mapview")
# attach QGIS output
# either do it "manually":
buf <- read_sf(out$OUTPUT[1])
# or use the st_as_sf.qgis_result method:
buf <- sf::st_as_sf(out)
# plot your result
mapview(buf, col.regions = "blue") + 
  mapview(random_points, col.regions = "red", cex = 3)
```

You can convert each qgis-algorithm into an R function with `qgis_function()`.
So using our buffer example from above, we could also run:

```{r function-creation, eval=FALSE}
# create a function
qgis_buffer <- qgis_function("native:buffer")
# run the function
out <- qgis_buffer(INPUT = random_points, DISTANCE = 50)
```

## Second example

As a second example, let's have a look at how to do raster processing running GRASS in the background.
To compute various terrain attributes of a digital elevation model, we can use `grass7:r.slope.aspect`.

`qgis_description()` (also included in `qgis_show_help()`) gives us the general description of the algorithm.

```{r desc}
qgis_description("grass7:r.slope.aspect")
```

We can find out about the arguments again with the help of `qgis_arguments()`.

```{r args}
qgis_arguments("grass7:r.slope.aspect")
```

`qgis_outputs()` shows the different outputs that will be calculated:

```{r outputs}
qgis_outputs("grass7:r.slope.aspect")
```

Now let us calculate the terrain attributes.

```{r slopeaspect, message=FALSE}
library("terra")
# attach digital elevation model from Mt. Mongón (Peru)
dem <- rast(system.file("raster/dem.tif", package = "spDataLarge"))

info <- qgis_run_algorithm(alg = "grass7:r.slope.aspect", elevation = dem)
                           
```

Just printing the `info` object shows which output objects have been made:

```{r info}
info
```

Combine these output rasters as a multi-layered `SpatRaster` object and plot it:

```{r combine1, message=FALSE}
# just keep the names of output rasters
nms <- qgis_outputs("grass7:r.slope.aspect")$name
# read in the output rasters 
r <- info[nms] |>
  unlist() |>
  terra::rast() |>
  as.numeric()
names(r) <- nms
# plot the output
plot(r)
```

An alternative way to combine the rasters is given below.

```{r combine2, message=FALSE}
r <- lapply(info[nms], \(x) as.numeric(qgis_as_terra(x))) |> 
  rast()
plot(r)
```
