---
title: "1. Basic usage of qgisprocess"
author: "Jannes Muenchow"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{1. Basic usage of qgisprocess}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

# Setting up the system

**qgisprocess** is basically a wrapper around the standalone command line tool [`qgis_process`](https://github.com/qgis/QGIS-Documentation/issues/5803).
Therefore, you need to have installed QGIS on your system as well as 3rd party providers such as GRASS and SAGA to access and run all geoalgorithms provided through `qgis_process` from within R.
To facilitate using **qgisprocess**, we have created a docker image that already comes with the needed software packages.
You can pull it from dockerhub by running:

```{sh, eval=FALSE}
docker pull geocompr/geocompr:qgis-ext
```

For a more detailed introduction on how to use docker with **qgisprocess** have a look a the corresponding vignette.

# How to use qgisprocess

Since **qgisprocess** depends on the command line tool `qgis_process`, it already tries to detect `qgis_process` on your system when it is being attached, and complains if it cannot find it.

```{r}
library("qgisprocess")
```

If you want to find out, how **qgisprocess**  detected `qgis_process`, run `qgis_configure()`.

```{r config}
qgis_configure()
```

Here, `qgis_process` was detected in the PATH variable.
However, `qgis_configure()` first tries if the R option `qgisprocess.path` or the global environment variable `R_QGISPROCESS_PATH` has been set.
This already indicates that you can specify one of these variables in case `qgis_process` has not been installed in one of the most common locations or if there are multiple QGIS versions available (both of which might sometimes happen under Windows).
If this is the case, set `options(qgisprocess.path = '/path/to/qgis_process')` and re-run `qgis_configure()`.
Under Windows make sure to indicate the path to the `qgis_process-qgis.bat` file, e.g.,

```{r win-config, eval=FALSE}
# specify path to QGIS installation on Windows 
options("qgisprocess.path" = "C:/Program Files/QGIS 3.16/bin/qgis_process-qgis.bat")
# or use the QGIS nightly version (if installed via OSGeo4W)
# options("qgisprocess.path" = "C:/OSGeo4W64/bin/qgis_process-qgis-dev.bat")
qgis_configure()
```

Assuming that `qgis_configure()` ran successfully, we can check which QGIS version our system is running:

```{r vers}
qgis_version()
```

Next, let's list all available providers including available 3rd party applications: 

```{r providers}
qgis_providers()
```

This tells us that we can also use the third-party providers GDAL, SAGA and GRASS trough the QGIS interface.
To find out about the available geoalgorithms, run:

```{r algs}
algs = qgis_algorithms()
algs
# or use the cached algorithms which have been collected while
# calling library(qgisprocess) under Windows
# algs = qgisprocess_cache$algorithms
```

Since we have also installed GRASS and SAGA, nearly 1000 geoalgorithms are at our disposal.
To find out about a specific geoalgorithm and a description of its arguments, use `qgis_show_help()`, e.g.:

```{r desc}
qgis_show_help("native:buffer")
# cat(qgis_description("native:buffer"))
# cat(qgis_help_text("native:buffer"))
```

Find out the arguments of a specific geoalgorithm:

```{r args}
qgis_arguments("native:buffer")
```


Vector example

If you would like to have an R-like experience running QGIS, create R functions first.

```{r function-creation}
qgis_buffer = qgis_function("native:buffer")
```


Raster example (GRASS7)

```{r}
qgis_arguments("grass7:r.slope.aspect")
```

```{r slopeaspect}
library("raster")
# attach digital elevation model from Mt. Mongón (Peru)
data("dem", package = "spDataLarge")
projection(dem) = sf::st_crs(32717)$proj4string

info = qgis_run_algorithm(alg = "grass7:r.slope.aspect",
                          elevation = dem, 
                          .quiet = TRUE)
# just keep the names of output rasters
nms = grep("^\\.", names(info), invert = TRUE, value = TRUE)
# read in the output rasters as a raster stack
s = stack(lapply(info[nms], unclass))
# plot the output
plot(s)

```

<!--
# raster example
## ----slope_aspect, eval = TRUE, messages = FALSE, cache = FALSE----------
params$elevation <- dem
params$pcurvature <- file.path(tempdir(), "pcurv.tif")
params$tcurvature <- file.path(tempdir(), "tcurv.tif")
out <- run_qgis(alg = "grass7:r.slope.aspect",
                params = params,
                load_output = TRUE,
                show_output_paths = FALSE)
class(out)
names(out)

## ----eval = FALSE--------------------------------------------------------
library("raster")
plot(stack(out))

-->
