---
title: "1. Basic usage of qgisprocess"
author: "Jannes Muenchow"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{1. Basic usage of qgisprocess}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

**qgisprocess** is basically a wrapper around the standalone command line tool [`qgis_process`](https://github.com/qgis/QGIS-Documentation/issues/5803).
Therefore, it already tries to detect where `qgis_process` is installed on your system when it is being attached.

```{r}
library("qgisprocess")
```

If you want to find out, how **qgisprocess**  detected `qgis_process`, run `qgis_configure()`.

```{r config}
qgis_configure()
```

Here, `qgis_process` was detected in the PATH variable.
However, `qgis_configure()` first tries if the R option `qgisprocess.path` or the global environment variable `R_QGISPROCESS_PATH` has been set.
This already indicates that you can specify one of these variables in case `qgis_process` has not been installed in one of the most common locations or if there are multiple QGIS versions available both of which might sometimes happen under Windows.
If this is th case, set `options(qgisprocess.path = '/path/to/qgis_process')` and re-run `qgis_configure()`.
Under Windows make sure to indicate the path to the `qgis_process-qgis.bat` file, e.g.,

```{r win-config, eval=FALSE}
# specify path to QGIS installation on Windows 
options("qgisprocess.path" = "C:/Program Files/QGIS 3.16/bin/qgis_process-qgis.bat")
options(qgisprocess.path = "C:/OSGeo4W64/bin/qgis_process-qgis-dev.bat")
qgis_configure()
```

Assuming that `qgis_configure()` ran successfully, we can check which QGIS version our system is running:

```{r vers}
qgis_version()
```

Next, let's list all available providers including available 3rd party applications: 

```{r providers}
qgis_providers()
```

List all geoalgorithms at your disposal:

```{r algs}
qgis_algorithms()
# or use the cached algorithms which have been collected while
# calling library(qgisprocess) under Windows
# algs = qgisprocess_cache$algorithms
```

Find out what a specific geoalgorithm is doing, e.g., `native:buffer`.

```{r desc}
qgis_show_help("native:buffer")
# cat(qgis_description("native:buffer"))
# cat(qgis_help_text("native:buffer"))
```

Find out the arguments of a specific geoalgorithm:

```{r args}
qgis_arguments("native:buffer")
```


Vector example

If you would like to have an R-like experience running QGIS, create R functions first.

```{r function-creation}
qgis_buffer = qgis_function("native:buffer")
```


Raster example (GRASS7)

```{r}
qgis_arguments("grass7:r.slope.aspect")
```

```{r slopeaspect}
library("raster")
data("dem", package = "spDataLarge")
projection(dem) = sf::st_crs(32717)$proj4string

# works under QGIS 3.16.1 but not under 3.17 (nightly version)
info = qgis_run_algorithm(alg = "grass7:r.slope.aspect",
                elevation = dem, 
                # pcurvature = file.path(tempdir(), "pcurv.tif"),
                # tcurvature = file.path(tempdir(), "tcurv.tif"),
                .quiet = FALSE)
raster(qgis_output(info, "pcurvature")[1])
```

<!--
# raster example
## ----slope_aspect, eval = TRUE, messages = FALSE, cache = FALSE----------
params$elevation <- dem
params$pcurvature <- file.path(tempdir(), "pcurv.tif")
params$tcurvature <- file.path(tempdir(), "tcurv.tif")
out <- run_qgis(alg = "grass7:r.slope.aspect",
                params = params,
                load_output = TRUE,
                show_output_paths = FALSE)
class(out)
names(out)

## ----eval = FALSE--------------------------------------------------------
library("raster")
plot(stack(out))

-->
