```{r setup, message=FALSE}
library(qgisprocess)
```

When you run `qgis_show_help()` or `qgis_get_argument_specs()` for a given algorithm, you will quickly find out that QGIS has a diverse set of possible argument types.

As an example, take a look at the `qgis_type` column below:

```{r}
qgis_get_argument_specs("native:joinbynearest") |> 
  subset(select = name:qgis_type)
```

This article addresses the question how algorithm arguments in `qgis_run_algorithm()` should be formatted.

Although you can pass a string [^string] value in the way QGIS expects it, `{qgisprocess}` makes it possible to pass familiar R objects that naturally match the QGIS argument type.
This is often easier than constructing a string in the format required by QGIS.

[^string]: A string is a character vector of length 1.

The tables below show which R objects can be passed to each QGIS argument type!
You can find a bit more background information in issue [#13](https://github.com/r-spatial/qgisprocess/issues/13).

Note that some arguments will receive a default value if they are unspecified.
Defaults are provided by either `{qgisprocess}` or QGIS.
Typical defaults by `{qgisprocess}` are temporary filepaths for (missing) output argument types `sink`, `vectorDestination`, `rasterDestination`, `fileDestination` and `folderDestination`.

# Spatial QGIS arguments

QGIS argument type | Supported R object
-- | -----
`source`, `vector` | a string (filepath) or an appropriate spatial vector object (`sf`, `SpatVector`, `SpatVectorProxy`)
`raster` | a string (filepath) or an appropriate spatial raster object (`SpatRaster`, `stars`, `stars_proxy`, `RasterLayer`, `RasterBrick`)
`layer` |  a string (filepath) or an appropriate spatial object (`sf`, `SpatVector`, `SpatVectorProxy`, `SpatRaster`, `stars`, `stars_proxy`, `RasterLayer`, `RasterBrick`)
`multilayer` | a list of layers created by `qgis_list_input()` (alternatively, repeat the same argument providing each layer in turn)
`aggregates`, `field_mapping`, `tininputlayers`, `vectortilewriterlayers` | a nested list created by `qgis_list_input()` (unnamed list) or `qgis_dict_input()` (named list) or a combination of both (a `native:aggregate` example is found [here](https://github.com/r-spatial/qgisprocess/issues/133#issuecomment-1488490056)). These argument types are typically not supported by the legacy (no-JSON) input method.
`point` | a vector of 2 point coordinates, a simple feature geometry (`sfg`) of class `POINT`, or an `sfc` (geometry set) or `sf` object with exactly one `POINT` geometry
`band` | an integer value
`extent` | a vector of the form `c(xmin, xmax, ymin, ymax)`, a `bbox` object from `{sf}`, a `SpatExtent` object from `{terra}` or an `Extent` object from `{raster}`
`crs` | a `crs` object from `{sf}`, a `CRS` object from `{raster}` or a WKT2 string (e.g. obtained with `terra::crs()`)
`coordinateoperation` | PROJ string of a coordinate operation, possibly obtained using `sf::sf_proj_pipelines()`
`sink`, `vectorDestination` | a string: filepath to a vector file format (defaults to a temporary GeoPackage if argument is missing)
`rasterDestination` | a string: filepath to a raster file format (defaults to a temporary GeoTIFF file if argument is missing)

# Non-spatial QGIS arguments

QGIS argument types | Supported R object
-- | -----
`enum` | a character vector with one or more acceptable string values. A numeric vector with the corresponding index number(s) can also be provided, but contrary to the character vector its acceptability won't be checked.
`range` | a vector of length 2, defining minimum and maximum value respectively
`file`, `field`, `layout`, `layoutitem`, `maptheme`, `execute_sql` | string
`string` | any string, including data-defined overriding (`"field:..."` or `"expression:..."`)
`distance`, `number` | numeric (length 1), or a string for data-defined overriding (`"field:..."` or `"expression:..."`)
`boolean` | logical (length 1), or a string for data-defined overriding (`"field:..."` or `"expression:..."`)
`color` | a colour string that `col2rgb()` understands (e.g. `"pink1"` or `"#1A664D80"`), or a string for data-defined overriding (`"field:..."` or `"expression:..."`)
`expression`, `raster_calc_expression` | a string (formatted as `"expression:..."`). See `vignette("qgis_expressions")`.
`matrix` | a matrix or a data frame with contents as required by the algorithm
`relief_colors` | a matrix or a data frame with three columns and with rows corresponding to intervals. The first two columns define the interval (minimum and maximum respectively) and the third column must be a colour string that `col2rgb()` understands (e.g. `"pink1"` or `"#1A664D80"`).
`fileDestination`, `folderDestination` | a string: path to a file or directory (defaults to a temporary file or directory if argument is missing)


