For a very short introduction to **qgisprocess**, visit the [homepage](https://r-spatial.github.io/qgisprocess).

Here you will learn about package configuration, about basic usage illustrated by two examples, and how to pipe results into a next geoprocessing step.

# Setting up the system

**qgisprocess** is basically a wrapper around the standalone command line tool [`qgis_process`](https://docs.qgis.org/latest/en/docs/user_manual/processing/standalone.html).
Therefore, you need to have installed QGIS on your system as well as third-party providers such as GRASS and SAGA to access and run all geoalgorithms provided through `qgis_process` from within R.

The package is meant to support _current_ QGIS releases, i.e. both the latest and the long-term release.
Although older QGIS releases are not officially supported, **qgisprocess** might work with QGIS versions >=3.16.
Download instructions for all platforms are available at <https://download.qgis.org/>.

To facilitate using **qgisprocess**, we have created a docker image that already comes with the needed software packages.
You can pull it from Github's container registry by running:

```{sh, eval=FALSE}
docker pull ghcr.io/geocompx/docker:qgis
```

For a more detailed introduction on how to get started with docker, please refer to https://github.com/geocompx/docker.

# Package configuration

Since **qgisprocess** depends on the command line tool `qgis_process`, it already tries to detect `qgis_process` on your system when it is being loaded, and complains if it cannot find it.

```{r}
library("qgisprocess")
```

When loading **qgisprocess** for the first time, it will cache among others the path to `qgis_process`, the QGIS version and the list of known algorithms.
When loading **qgisprocess** in later R sessions, the cache file is read instead for speed-up, on condition that it is still valid.
Therefore, usually you don't have to do any configuration yourself, unless there's a message telling you to do so.

If you are interested in the details about this process, e.g. how **qgisprocess**  detected `qgis_process`, run `qgis_configure(use_cached_data = TRUE)`.

```{r config}
qgis_configure(use_cached_data = TRUE)
```

If needed the cache will be rebuilt automatically upon loading the package.
This is the case when the QGIS version or the location of the `qgis_process` command-line utility has changed, user-settings (e.g. the option `qgisprocess.path`) have been altered or a changed state of the processing provider plugins (enabled vs. disabled) has been detected.

Rebuilding the cache can be triggered manually by running `qgis_configure()` (its default is `use_cached_data = FALSE`).

To determine the location of `qgis_process`, `qgis_configure()` first checks if the R option `qgisprocess.path` or the global environment variable `R_QGISPROCESS_PATH` has been set.
This already indicates that you can specify one of these settings in case `qgis_process` has not been installed in one of the most common locations or if there are multiple QGIS versions available.
If this is the case, set `options(qgisprocess.path = '/path/to/qgis_process')` or set the environment variable (e.g. in `.Renviron`) and run `qgis_configure()`.
Under Windows make sure to indicate the path to the `qgis_process-qgis.bat` file, e.g.,

```{r win-config, eval=FALSE}
# specify path to QGIS installation on Windows 
options(qgisprocess.path = "C:/Program Files/QGIS 3.28/bin/qgis_process-qgis.bat")
# or use the QGIS nightly version (if installed via OSGeo4W)
# options(qgisprocess.path = "C:/OSGeo4W64/bin/qgis_process-qgis-dev.bat")
qgis_configure() # or use library(qgisprocess) if package was not loaded yet
```

Assuming that package loading or `qgis_configure()` ran successfully, we can check which QGIS version our system is running (it takes this from the cache):

```{r vers}
qgis_version()
```

Next, let's check which plugins are at our disposal:

```{r plugins}
qgis_plugins()
```

Since we will use GRASS and SAGA later on, let's enable them (of course, this implies that you have already installed GRASS and SAGA version > 7 on your system):

```{r enable-plugins, message=FALSE}
qgis_enable_plugins(c("grassprovider", "processing_saga_nextgen"))
```

Now, let's list all available providers including available third-party applications: 

```{r providers}
qgis_providers()
```

This tells us that we can also use the third-party providers GDAL, GRASS and SAGA through the QGIS interface.

# Basic usage

## First example

To find out about the available (cached) geoalgorithms, run:

```{r algs}
algs <- qgis_algorithms()
algs
```

Since we have also installed GRASS and SAGA, nearly 1000 geoalgorithms are at our disposal.
To find out about a specific geoalgorithm and a description of its arguments, use `qgis_show_help()`, e.g.:

```{r help, eval=FALSE}
qgis_show_help("native:buffer")
## Buffer (native:buffer)
## 
## ----------------
## Description
## ----------------
## This algorithm computes a buffer area for all the features in an input layer, using a fixed or dynamic distance.
## 
## The segments parameter controls the number of line segments to use to approximate a quarter circle when creating rounded offsets.
## 
##...
```

To find out the arguments of a specific geoalgorithm, run:

```{r args-buffer}
qgis_get_argument_specs("native:buffer")
```

And finally run it with `qgis_run_algorithm()`:

```{r buffer}
# if needed, first install spDataLarge:
# remotes::install_github("Nowosad/spDataLarge")
data("random_points", package = "spDataLarge")
result <- qgis_run_algorithm("native:buffer", INPUT = random_points, DISTANCE = 50)
```

As a convenience to the user, `qgis_run_algorithm()` reports all unspecified and automatically chosen arguments.
If you want to have even more information on what is going on in the background, set `.quiet` to `FALSE`.
The `result` object is of class `qgis_result` and contains the path to the output file created by `qgis_process` (when not explicitly setting an output filepath, **qgisprocess** creates it automatically for you).
The output filepath can be extracted with `qgis_extract_output()`.
`qgis_result` objects are of type `list` which, aside from the geoprocessing result, also contain debugging information about the used algorithm, input arguments and messages from the processing step.
See `?qgis_result_status` for various convenience functions to extract all of this information easily from `qgis_result` objects.

For QGIS 3.24 and later, `qgis_run_algorithm()` passes the input arguments to QGIS as a JSON string.
The JSON input string is also included in `qgis_result` objects.
Moreover, the user can specify input arguments directly as JSON in `qgis_run_algorithm()`.
That is useful since input parameters can be copied from the QGIS GUI as JSON.
This will be demonstrated in a separate tutorial.

```{r inspect}
# inspect the result object
class(result)
names(result)
result # only prints the output element(s)
```

To read in the QGIS output and visualize it, we can run:

```{r plot-buffer, message=FALSE}
library("sf")
library("mapview")
# attach QGIS output
# either do it "manually":
buf <- read_sf(qgis_extract_output(result, "OUTPUT"))
# or use the st_as_sf.qgis_result method:
buf <- sf::st_as_sf(result)
# plot your result
mapview(buf, col.regions = "blue") + 
  mapview(random_points, col.regions = "red", cex = 3)
```

You can convert each qgis-algorithm into an R function with `qgis_function()`.
So using our buffer example from above, we could also run:

```{r function-creation, eval=FALSE}
# create a function
qgis_buffer <- qgis_function("native:buffer")
# run the function
result <- qgis_buffer(INPUT = random_points, DISTANCE = 50)
```

This is basically what package [qgis](https://github.com/JanCaha/r_package_qgis) is doing for each available QGIS function while also providing an R help file for each function.
Hence, if you prefer running QGIS with callable R functions, check it out.


## Second example

As a second example, let's have a look at how to do raster processing running GRASS in the background.
To compute various terrain attributes of a digital elevation model, we can use `grass7:r.slope.aspect`.

`qgis_get_description()` (also included in `qgis_show_help()`) gives us the general description of the algorithm.

```{r desc}
qgis_get_description("grass7:r.slope.aspect")
```

We can find out about the arguments again with the help of `qgis_get_argument_specs()`.

```{r args}
qgis_get_argument_specs("grass7:r.slope.aspect")
```

`qgis_get_output_specs()` shows the different outputs that will be calculated:

```{r outputs}
qgis_get_output_specs("grass7:r.slope.aspect")
```

Now let us calculate the terrain attributes.

```{r slopeaspect, message=FALSE}
library("terra")
# attach digital elevation model from Mt. Mongón (Peru)
dem <- rast(system.file("raster/dem.tif", package = "spDataLarge"))
# if not already done, enable the grass plugin
# qgis_enable_plugins("grassprovider")
info <- qgis_run_algorithm(alg = "grass7:r.slope.aspect", elevation = dem)
```

Just printing the `info` object shows which output files have been made:

```{r info}
info
```

Combine these output rasters as a multi-layered `SpatRaster` object and plot it:

```{r combine1, message=FALSE}
# just keep the names of output rasters
nms <- qgis_get_output_specs("grass7:r.slope.aspect")$name
# read in the output rasters 
r <- info[nms] |>
  unlist() |>
  terra::rast() |>
  as.numeric()
names(r) <- nms
# plot the output
plot(r)
```

An alternative way to combine the rasters is given below.

```{r combine2, message=FALSE, eval=FALSE}
r <- lapply(info[nms], \(x) as.numeric(qgis_as_terra(x))) |> 
  rast()
```

Since we have now many terrain attributes at our disposal, let us take the opportunity to add their values to points laying on top of them with the help of the SAGA function `sagang:addrastervaluestopoints`.

```{r addrastertopoints-args}
qgis_get_argument_specs("sagang:addrastervaluestopoints")
```

The `GRIDS` argument is of type multilayer.
To pass multiple layers to one argument, you can either repeat the corresponding argument as often as needed ... 

```{r multilayer-repeat}
rp_tp <- qgis_run_algorithm(
  "sagang:addrastervaluestopoints",
  SHAPES = random_points,
  GRIDS = info$aspect[[1]],
  GRIDS = info$slope[[1]],
  GRIDS = info$tcurvature[[1]],
  RESAMPLING = 0)
```

... or you can pass to it all needed layers in one list.
One could use the `list()` command but it is recommendended to use the `qgis_list_input()` function which is more robust, and therefore will also support non-JSON-input configurations (e.g. QGIS < 3.24).

```{r multilayer-list, eval=FALSE}
rp_tp <- qgis_run_algorithm(
  "sagang:addrastervaluestopoints",
  SHAPES = random_points,
  GRIDS = qgis_list_input(info$aspect[[1]], info$slope[[1]], info$tcurvature[[1]]),
  RESAMPLING = 0)
```

To verify that it worked, read in the output.

```{r multilayer-output}
sf::st_as_sf(rp_tp)
```
# Piping

`qgis_process` does not lend itself naturally to piping because its first argument is the name of a geoalgorithm instead of a data object.
`qgis_run_algorithm_p()` circumvents this by accepting a `.data` object as its first argument, and pipes this data object into the the first argument of a geoalgorithm assuming that the specified geoalgorithm needs a data input object as its first argument.

```{r buf-pipe}
system.file("longlake/longlake_depth.gpkg", package = "qgisprocess") |>
  qgis_run_algorithm_p("native:buffer", DISTANCE = 50)
```

If `.data` is a `qgis_result` object, `qgis_run_algorithm_p()` automatically tries to select an element named `OUTPUT`.
However, if the output has another name (e.g., `DEM_PREPROC` as in the example below) or if there are multiple output elements to choose from (e.g., `sagang:sagawetnessindex` has four output rasters, check with `qgis_outputs("sagang:sagawetnessindex")`), you can specify the wanted output object via the `.select` argument.
Please note that we make sure that temporary output raster files, i.e., all output rasters we do not specifically name ourselves, should use SAGA's native raster file format by setting the `qgisprocess.tmp_raster_ext` to `.sdat`.
Using the default raster output format `.tif` might lead to trouble depending on the installed versions of third-party packages (GDAL, SAGA, etc.). 

```{r qgis-pipe}
dem <- system.file("raster/dem.tif", package = "spDataLarge")
# in case you need to enable the SAGA next generation algorithms, run the following line:
# qgis_enable_plugins("processing_saga_nextgen")

options(qgisprocess.tmp_raster_ext = ".sdat")
qgis_run_algorithm(algorithm = "sagang:sinkremoval", DEM = dem,
                   METHOD = 1) |>
  qgis_run_algorithm_p("sagang:sagawetnessindex", .select = "DEM_PREPROC")
```

When piping, `qgis_run_algorithm_p()` automatically cleans up after you by deleting intermediate results.
This avoids cluttering your system when running geoalgorithms on large spatial data files.
To turn off this behavior, set `.clean` to `FALSE`.

Of course, you can also pipe to `qgis_run_algorithm()` by manually extracting the `OUTPUT` object and redirecting it to the appropriate input argument of the next processing step.
This avoids ambiguity and allows for greater flexibility though it might not be as convenient as `qgis_run_algorithm_p()`.
For example, intermediate results remain on disk for the duration of your R session, unless you manually call `qgis_clean_result()` on a result object.

```{r native pipe}
result <- qgis_run_algorithm(algorithm = "sagang:sinkremoval", DEM = dem, 
                         METHOD = 1) |>
  _$DEM_PREPROC[1] |>
  qgis_run_algorithm(algorithm = "sagang:sagawetnessindex",
                     DEM = _)

# or using an anonymous function
# result <- qgis_run_algorithm(algorithm = "sagang:sinkremoval", DEM = dem, 
#                          METHOD = 1) |>
#   (\(x) qgis_run_algorithm(algorithm = "sagang:sagawetnessindex",
#                            DEM = x$DEM_PREPROC[1])) ()

# set the default output raster format to .tif again
options(qgisprocess.tmp_raster_ext = ".tif")
```
